---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EpiGraphs

The goal of EpiGraphs is to allow for epidemiologists and statisticians to graph data quick, easy, and in a professional manner.

## Installation

You can install the development version of EpiGraphs from https://github.com/chaseramirez/EpiGraphs or devtools::install_github("chaseramirez/EpiGraphs")

``` r
# install.packages("pak")
pak::pak("chaseramirez/EpiGraphs")
```

## Example

This is a basic example which shows you how to solve a common problem:
Using congenital syphilis data, download from:
https://www.who.int/data/gho/data/indicators/indicator-details/GHO/congenital-syphilis-number-of-reported-cases
Click on the data option and follow the instructions on the right side of the screen to export the data. Do not add any filters.

```{r example}
syphilis <- read.csv("C:/Users/crami/OneDrive/Desktop/Undergrad Class Folders/Fall 2024 B481 - Biostats Computing/Syphilis.csv")

library(dplyr)
library(ggplot2)
library(scales)
library(EpiGraphs)

# Example: Top 10 locations with the highest syphilis cases in 2022
epi_bar_plot(
  data = syphilis,          # Your dataset
  x = "Location",           # Categorical variable for the x-axis
  y = "Value",              # Numeric variable for bar heights
  period_var = "Period",      # Period column (years, months, ect.)
  period = 2022,              # Filter for the year 2022
  top_n = 10,               # Display top 10 locations
  disease_name = "Cong. Syphilis", # Disease name for the title
  y_desc = "Cases"          # Description for the y-axis (e.g., "cases"), synced with title
)

```
