---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EpiGraphs

<!-- badges: start -->
<!-- badges: end -->

The goal of EpiGraphs is to create simpler ways for epidemiologists and statiticians to visualize and analyze data.

## Installation

You can install the development version of EpiGraphs from https://github.com/chaseramirez/EpiGraphs.git.

``` r
# install.packages("pak")
pak::pak("chaseramirez/EpiGraphs")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(EpiGraphs)
# Example: Top 10 locations with the highest syphilis cases in 2022
epi_bar_plot(
  data = syphilis,          # Your dataset
  x = "Location",           # Categorical variable for the x-axis
  y = "Value",              # Numeric variable for bar heights
  period_var = "Period",      # Period column (years, months, ect.)
  period = 2022,              # Filter for the year 2022
  top_n = 10,               # Display top 10 locations
  disease_name = "Cong. Syphilis", # Disease name for the title
  y_desc = "Cases"          # Description for the y-axis (e.g., "cases"), synced with title
)


